Edge sign prediction in general graphs and Correlation Clustering

### 2017-09-20 troll/diff_mcc_pp.tex
mark content added from the aistats paper
minor corrections in troll experiments and add missing figures

\begin{table}[htpb]
  \centering
  \caption{100 time the difference of MCC between using the twin heuristic and not using it for our
  two main methods. Those values are computed on the same training/testing split as of
\autoref{tab:all_mcc}.\label{tab:twin}}
  \begin{tabular}{lrrrrrrr}
    \toprule
                                & $5\%$                    & $10\%$                   & $15\%$                   & $20\%$                   & $25\%$                   & $50\%$                   & $90\%$                   \\
    \midrule
    \multicolumn{8}{c}{\aut{}} \\
    \uslpropGsecTwin{}          & $0.0$                    & $0.0$                    & $0.0$                    & $0.0$                    & $0.0$                    & \textcolor{Red}{$-0.1$}  & \textcolor{Red}{$-0.7$}  \\
    \usruleTwin{}               & $0.0$                    & $0.0$                    & $0.0$                    & $0.0$                    & \textcolor{Green}{$0.1$} & \textcolor{Red}{$-0.2$}  & \textcolor{Red}{$-0.7$}  \\
    \multicolumn{8}{c}{\adv{}} \\
    \uslpropGsecTwin{}          & \textcolor{Red}{$-0.2$}  & \textcolor{Red}{$-0.6$}  & \textcolor{Red}{$-1.0$}  & \textcolor{Red}{$-1.6$}  & \textcolor{Red}{$-2.2$}  & \textcolor{Red}{$-5.4$}  & \textcolor{Red}{$-10.8$} \\
    \usruleTwin{}               & \textcolor{Red}{$-0.1$}  & \textcolor{Red}{$-0.5$}  & \textcolor{Red}{$-1.0$}  & \textcolor{Red}{$-1.6$}  & \textcolor{Red}{$-2.2$}  & \textcolor{Red}{$-5.7$}  & \textcolor{Red}{$-11.1$} \\
    \multicolumn{8}{c}{\wik{}} \\
    \uslpropGsecTwin{}          & $0.0$                    & $0.0$                    & $0.0$                    & $0.0$                    & $0.0$                    & \textcolor{Red}{$-0.1$}  & \textcolor{Red}{$-0.2$}  \\
    \usruleTwin{}               & $0.0$                    & $0.0$                    & $0.0$                    & \textcolor{Red}{$-0.1$}  & \textcolor{Red}{$-0.1$}  & \textcolor{Red}{$-0.2$}  & \textcolor{Red}{$-0.2$}  \\
    \multicolumn{8}{c}{\sla{}} \\
    \uslpropGsecTwin{}          & \textcolor{Green}{$0.3$} & \textcolor{Green}{$0.5$} & \textcolor{Green}{$0.8$} & \textcolor{Green}{$0.9$} & \textcolor{Green}{$1.1$} & \textcolor{Green}{$2.0$} & \textcolor{Green}{$3.3$} \\
    \usruleTwin{}               & \textcolor{Green}{$0.3$} & \textcolor{Green}{$0.6$} & \textcolor{Green}{$0.9$} & \textcolor{Green}{$1.1$} & \textcolor{Green}{$1.3$} & \textcolor{Green}{$2.2$} & \textcolor{Green}{$3.5$} \\
    \multicolumn{8}{c}{\epi{}} \\
    \uslpropGsecTwin{}          & \textcolor{Green}{$0.3$} & \textcolor{Green}{$0.5$} & \textcolor{Green}{$0.7$} & \textcolor{Green}{$0.8$} & \textcolor{Green}{$1.0$} & \textcolor{Green}{$1.9$} & \textcolor{Green}{$3.0$} \\
    \usruleTwin{}               & \textcolor{Green}{$0.4$} & \textcolor{Green}{$0.6$} & \textcolor{Green}{$0.8$} & \textcolor{Green}{$1.0$} & \textcolor{Green}{$1.2$} & \textcolor{Green}{$1.8$} & \textcolor{Green}{$2.8$} \\
    \multicolumn{8}{c}{\kiw{}} \\
    \uslpropGsecTwin{}          & $0.0$                    & $0.0$                    & $0.0$                    & $0.0$                    & \textcolor{Red}{$-0.1$}  & \textcolor{Red}{$-0.1$}  & \textcolor{Red}{$-0.3$}  \\
    \usruleTwin{}               & $0.0$                    & $0.0$                    & $0.0$                    & $0.0$                    & $0.0$                    & \textcolor{Red}{$-0.2$}  & \textcolor{Red}{$-0.5$}  \\
    \bottomrule
  \end{tabular}
\end{table}

>>>>

% \begin{table}[htpb]
%   \centering
%   \caption{
\begin{center}
   \captionsetup{font=small}
   \captionof{table}{%
    100 time the difference of MCC between using the twin heuristic and not using it for our
  two main methods. Those values are computed on the same training/testing split as of
\autoref{tab:all_mcc}.\label{tab:twin}}
  \begin{tabular}{lrrrrrrr}
    \toprule
                                & $5\%$                    & $10\%$                   & $15\%$                   & $20\%$                   & $25\%$                   & $50\%$                   & $90\%$                   \\
    \midrule
    & \multicolumn{7}{c}{\aut{}} \\ \cmidrule(lr){2-8}
    \uslpropGsecTwin{}          & $0.0$                    & $0.0$                    & $0.0$                    & $0.0$                    & $0.0$                    & \textcolor{Red}{$-0.1$}  & \textcolor{Red}{$-0.7$}  \\
    \usruleTwin{}               & $0.0$                    & $0.0$                    & $0.0$                    & $0.0$                    & \textcolor{Green}{$0.1$} & \textcolor{Red}{$-0.2$}  & \textcolor{Red}{$-0.7$}  \\
    & \multicolumn{7}{c}{\adv{}} \\ \cmidrule(lr){2-8}
    \uslpropGsecTwin{}          & \textcolor{Red}{$-0.2$}  & \textcolor{Red}{$-0.6$}  & \textcolor{Red}{$-1.0$}  & \textcolor{Red}{$-1.6$}  & \textcolor{Red}{$-2.2$}  & \textcolor{Red}{$-5.4$}  & \textcolor{Red}{$-10.8$} \\
    \usruleTwin{}               & \textcolor{Red}{$-0.1$}  & \textcolor{Red}{$-0.5$}  & \textcolor{Red}{$-1.0$}  & \textcolor{Red}{$-1.6$}  & \textcolor{Red}{$-2.2$}  & \textcolor{Red}{$-5.7$}  & \textcolor{Red}{$-11.1$} \\
    & \multicolumn{7}{c}{\wik{}} \\ \cmidrule(lr){2-8}
    \uslpropGsecTwin{}          & $0.0$                    & $0.0$                    & $0.0$                    & $0.0$                    & $0.0$                    & \textcolor{Red}{$-0.1$}  & \textcolor{Red}{$-0.2$}  \\
    \usruleTwin{}               & $0.0$                    & $0.0$                    & $0.0$                    & \textcolor{Red}{$-0.1$}  & \textcolor{Red}{$-0.1$}  & \textcolor{Red}{$-0.2$}  & \textcolor{Red}{$-0.2$}  \\
    & \multicolumn{7}{c}{\sla{}} \\ \cmidrule(lr){2-8}
    \uslpropGsecTwin{}          & \textcolor{Green}{$0.3$} & \textcolor{Green}{$0.5$} & \textcolor{Green}{$0.8$} & \textcolor{Green}{$0.9$} & \textcolor{Green}{$1.1$} & \textcolor{Green}{$2.0$} & \textcolor{Green}{$3.3$} \\
    \usruleTwin{}               & \textcolor{Green}{$0.3$} & \textcolor{Green}{$0.6$} & \textcolor{Green}{$0.9$} & \textcolor{Green}{$1.1$} & \textcolor{Green}{$1.3$} & \textcolor{Green}{$2.2$} & \textcolor{Green}{$3.5$} \\
    & \multicolumn{7}{c}{\epi{}} \\ \cmidrule(lr){2-8}
    \uslpropGsecTwin{}          & \textcolor{Green}{$0.3$} & \textcolor{Green}{$0.5$} & \textcolor{Green}{$0.7$} & \textcolor{Green}{$0.8$} & \textcolor{Green}{$1.0$} & \textcolor{Green}{$1.9$} & \textcolor{Green}{$3.0$} \\
    \usruleTwin{}               & \textcolor{Green}{$0.4$} & \textcolor{Green}{$0.6$} & \textcolor{Green}{$0.8$} & \textcolor{Green}{$1.0$} & \textcolor{Green}{$1.2$} & \textcolor{Green}{$1.8$} & \textcolor{Green}{$2.8$} \\
    & \multicolumn{7}{c}{\kiw{}} \\ \cmidrule(lr){2-8}
    \uslpropGsecTwin{}          & $0.0$                    & $0.0$                    & $0.0$                    & $0.0$                    & \textcolor{Red}{$-0.1$}  & \textcolor{Red}{$-0.1$}  & \textcolor{Red}{$-0.3$}  \\
    \usruleTwin{}               & $0.0$                    & $0.0$                    & $0.0$                    & $0.0$                    & $0.0$                    & \textcolor{Red}{$-0.2$}  & \textcolor{Red}{$-0.5$}  \\
    \bottomrule
  \end{tabular}
% \end{table}
\end{center}

### 2017-09-20 troll/experiments.tex
minor corrections in troll experiments and add missing figures

We now evaluate our edge sign classification methods on representative real-world datasets of
varying density and label regularity.
After presenting the data and our evaluation criterion, we proceed in two steps. First, we simulate
our generative model on real networks to give them signs, then we study to which extent we can recover
the parameters $p$ and $q$ of each node, and how this translates to accurate prediction. Second, we
select random training sets from the actual signs.
This shows that our methods compete well against
existing approaches in terms of both predictive and computational performance. We are especially
interested in small training set regimes, and have restricted our comparison to the batch learning
scenario since all competing methods we are aware of have been developed in that setting only.

@@

We consider six real-world classification datasets. The first four are \dssn{} widely
used as benchmarks for this
task~(\eg{}\autocites{Leskovec2010}{shahriari2014ranking}{wu2016troll}{OnlineCompletion17}):
\begin{newcontent}
In \adv{}, a trust-based social network for open source developers, a user $u$ can certify another
user $v$ with different degrees of trust: \enquote{Observer}, \enquote{Apprentice} (both of which we
consider negative), \enquote{Journeyer} and \enquote{Master} (both of which we consider
positive).\footnote{We download the 7\thup{} of July, 2014 version from
\url{http://www.trustlet.org/datasets/advogato/}.} A full description of this trust metric and its
resistance to attacks are available in the PhD thesis of the website's creator~\autocite[Section
4]{AdvogatoTrustThesis02}).
\end{newcontent}
In \wik{}, there is an edge from user $u$ to user $v$ if $v$ applies for an admin position and $u$
votes for or against that promotion. In \sla{}, a news sharing and commenting website, member $u$
can tag other members $v$ as friends or foes. Finally, in \epi{}, an online shopping website, user
$v$ reviews products and, based on these reviews, another user $u$ can display whether he considers
$v$ to be reliable or not. In addition to these four datasets, we considered two other \ssn{} where
the signs are inferred automatically, rather than given explicitly by the users.  In
\kiw{}~\cite{wikiedits11}, an edge from Wikipedia user $u$ to user $v$ indicates whether they edited
the same article in a constructive manner or not.\footnote{This is the
\href{http://konect.uni-koblenz.de/networks/wikisigned-k2}{KONECT version of the
\enquote{Wikisigned} dataset}, from which we removed self-loops.} Finally, in the
\aut{}~\cite{kumar2016structure} network, an author $u$ cites another author $v$ by either endorsing
or criticizing $v$'s work. The edge sign is derived by classifying the citation sentiment with a
simple, yet powerful, keyword-based technique using a list of positive and negative words. See
\cite{kumar2016structure} for more details.\footnote{We again removed self-loops and merged
multi-edges which are all of the same sign.}

\begin{table*}[bt]
  \centering
  \small
  \caption{Dataset properties. The 5\thup{} column gives the fraction of positive labels. The next two
  columns provide two different measures of label regularity, while the last two columns give the
  proportion of reciprocal edges, and among them the fraction with different signs.\label{tab:dataset}}
  \begin{tabular}{lrrrrrrrr}
    \toprule
    Dataset & $|V|$       & $|E|$       & $\frac{|E|}{|V|}$ & $\frac{|E^+|}{|E|}$ & $\frac{\Psi^2_{G''}(Y)}{|E|}$ & $\frac{\Psi_G(Y)}{|E|}$ & twin edges & twin disagreement \\
    \midrule                                                                                                   
    \aut{}  & \np{4831}   & \np{39452}  & 8.2               & 72.3\%              & .076                          & .191                    & 5.1\%      & 27.1\%            \\
    \adv{}  & \np{5417}   & \np{51312}  & 9.5               & 75.1\%              & .061                          & .132                    & 33.6\%     & 28.6\%            \\
    \wik{}  & \np{7114}   & \np{103108} & 14.5              & 78.8\%              & .063                          & .142                    & 5.6\%      & 10.0\%            \\
    \sla{}  & \np{82140}  & \np{549202} & 6.7               & 77.4\%              & .059                          & .143                    & 17.7\%     & 4.0\%             \\
    \epi{}  & \np{131580} & \np{840799} & 6.4               & 85.3\%              & .031                          & .074                    & 30.8\%     & 2.1\%             \\
    \kiw{}  & \np{138587} & \np{740106} & 5.3               & 87.9\%              & .034                          & .086                    & 6.5\%      & 14.6\%            \\
    \bottomrule
  \end{tabular}
\end{table*}

\autoref{tab:dataset} summarizes statistics for these datasets. We note that most edge labels are
positive. Hence, test set accuracy is not an appropriate measure of prediction performance. We
instead evaluated our performance using the so-called Matthews Correlation Coefficient
(MCC)~\autocite{MCC00}), defined as
\[
	\mathrm{MCC} = \frac{tp\times tn-fp\times fn}%
        {\sqrt{ (tp + fp) ( tp + fn ) ( tn + fp ) ( tn + fn ) } } = \pm \sqrt{\frac{\chi^2}{n}}
\]
MCC combines all the four quantities found in a binary confusion matrix ($t$rue $p$ositive, $t$rue
$n$egative, $f$alse $p$ositive and $f$alse $n$egative) into a single metric which ranges from $-1$
(when all predictions are incorrect) to $+1$ (when all predictions are correct) through $0$ (when
predictions are made uniformly at random).

Although the semantics of the edge signs is not the same across these networks, we can see from
\autoref{tab:dataset} that our generative model essentially fits all of them. Specifically, the last
two columns of the table report the rate of label (ir)regularity, as measured by
$\Psi^2_{G''}(Y)/|E|$ (second-last column) and $\Psi_{G}(Y)/|E|$ (last column), where 
\[
\Psi^2_{G''}(Y) = \min_{(\bp,\bq)} \left(f_{E_0}(\bp,\bq) + f_{E\setminus E_0}(\bp,\bq)\right)\,,
\]
$f_{E_0}$ and $f_{E\setminus E_0}$ being the quadratic criterions of \autoref{ss:passive}, viewed as
functions of both $(\bp,\bq)$, and $y_{i,j}$ when all labels are known, and $\Psi_{G}(Y)$ is the
label regularity measure adopted in the online setting, as defined in \autoref{s:prel}. It is
reasonable to expect that higher label irregularity corresponds to lower prediction performance.
This trend is in fact confirmed by our experimental findings: whereas \epi{} tends to be easy,
\aut{} tends to be hard, and this holds for all algorithms we tested, even if they do not explicitly
comply with our inductive bias principles. Moreover, $\Psi^2_{G''}(Y)/|E|$ tends to be proportional
to $\Psi_{G}(Y)/|E|$ across datasets, hence confirming the anticipated connection between the two
regularity measures.

\begin{newcontent}
Finally, there is a low fraction of reciprocal (or \emph{twin}) edges (\ie{} both $\euv \in E$ and $\evu
\in E$), which is a common mechanism of link formation in directed
networks~\autocites{DirectedReciprocity04}{Reciprocity13}). Moreover, in most cases, such twin edges do not disagree,
\ie{} they have the same sign. In practice, we can use this fact to improve our accuracy at no
additional computational cost: when predicting $\euv \in E_{\mathrm{test}}$, if the reciprocal edge
\evu{} is part of the training set, we set $\yhat_{u,v} = y_{v,u}$. For clarity, when comparing
our methods with existing approaches, we not use that trick. But afterwards, we show in
\autoref{tab:twin} when it can be beneficial.
% more references in  http://www.nature.com/articles/srep02729
\end{newcontent}

@@

Recall that according to our generative model of \autoref{s:gen}, each node $u$ has parameter $p_u$
governing its sending behavior and each node $v$ has a parameter $q_v$ governing its receiving
behavior such that the sign of the edge \euv{} is positive with probability $\frac{p_u+q_v}{2} =
\etauv$. Given a topology $G=(V,E)$, we start by assigning a $p$ and $q$ values to each node. If we
want \etauv{} to be uniform over $E$, we have to take into account the out-degree of nodes with at
least one outgoing edges, and likewise the in-degree of nodes with at least one incoming edges. For
that, we partition the interval $[0,1]$ into $|V|$ segments of size proportional to $\dout(u_1),
\dout(u_2), \ldots, \dout(u_{|V|})$. We shuffle these segments and draw a number \uar{} from each of
them, which we set as the $p$ value of the corresponding node. Then, because we want to have more
positive edges, we apply an exponential transformation \[ p \rightarrow \frac{1}{1-e^{-\lambda}}
\left( 1-e^{-\lambda p} \right) \,,\] where we choose $\lambda=3$. We do the same for $q$, then for
every edge \euv{}, we set it positive with probability \etauv{}. An example of the distributions we
obtain for the \wik{} graph is showed on \autoref{fig:troll_synth_distrib}.

Once the signs are generated, we select a training set \trainset{} \uar{}, and we want to predict
the signs of the remaining edges in the testing set $E\setminus \trainset{}$. The Bayes optimal
predictor classifies an edge \euv{} as positive if $\etauv \geq \shalf$ and negative otherwise. 
We compare it with the algorithm analyzed in \autoref{ss:bayes_approx}, which we call \usrule{}
(Bayes Learning Classifier based on \emph{tr}ollness and \emph{un}trustworthiness). In practice,
\usrule{} proceeds as follow: after computing $\htr(u)$ and $\hun(u)$ on training set $E_0$ for all
$u \in V$ (or setting those values to \shalf{} in case there is no outgoing or incoming edges for
some node), we use the equation \eqref{eq:predictor}, having estimated $\tau$ on $E_0$.

The results of this comparison on the six datasets are showed in \autoref{tab:troll_synth_global}.
For each network, we generated the signs once. Then for different training size (20\%, 40\% or 80\%
of $E$), we sampled a training set and predict using the knowledge of the true $\eta$ (Bayes
predictor) or its estimated value (\usrule{}).
The MCC of the Bayes predictor is the same on all datasets (around $30$), and so is its accuracy
(around $75$). The corresponding values for \usrule{} are close on all datasets (although some are
easier) and the gap naturally decreases as the training size increases. Another interesting
quantity is how close can \usrule{} estimates $\eta$ using $\wh{\eta} = \left(1-\htr(u)\right) +
\left(1-\hun(v)\right) -\tauhat$ as defined in \autoref{ss:bayes_approx}. We compute the mean
absolute error (MAE) between $\eta$ and $\wh{\eta}$ for all testing edges, and specifically for
the testing edges whose both endpoints have been sampled above a certain threshold in the training
set. As expected, the estimation is more accurate with increasing training set size, and with
increasing number of samples for a given edge.

@@

Finally, we can also look in more details at the role of $p$ and $q$ by building a 2D histogram of
the testing edges \euv{} based on their coordinate $(p_u, q_v)$. For instance on
\autoref{fig:tr_synth_mae}, we see that the MAE between $\eta$ and $\wh{\eta}$ is not uniform over
the $p,q$-space. It seems to decrease along the diagonal from $(0,0)$ to $(1,1)$. Note though that
\autoref{fig:tr_synth_sizemae} suggests this is likely due to the higher number of edges in the top
right corner (reflecting the imbalance toward positive edges), which allows better estimation.

In this case, we cannot use the MCC because off-diagonal, the Bayes predictor classifies edges
either all positive or all negative, resulting in a division by zero in the definition. We thus
fall back on accuracy and display in \autoref{fig:tr_synth_acc} the difference between the accuracy
of the Bayes predictor and the accuracy of \usrule{}. This time, the anti diagonal where
$\frac{p+q}{2}=\lhalf$ seems to play a special role. Indeed, as showed more clearly in
\autoref{fig:tr_synth_cmpacc}, the gap in accuracy between the Bayes predictor and \usrule{}
increases symmetrically as $\frac{p+q}{2}$ moves away from \shalf{}. The gap also tends to disappear
when $\frac{p+q}{2}=\lhalf$, most likely because in that area, prediction can only be made by
chance.

While we show those patterns on \wik{} with a 40\% training size, they are consistent across
datasets and training size.

@@

We perform two other sets of experiments. The first one assesses the effect of predicting reciprocal
edges by their value if available. As expected, the results in \autoref{tab:twin} demonstrate that
it improves MCC when the network has enough twin edges with low disagreements, like in \sla{} and
\epi{}. It has no effect when there are few twin edges, like in \aut{}, \wik{} and \kiw{}. And it is
detrimental when there are many twin edges that do not agree enough, like in \adv{}, in which
case it is better to rely solely on the learned model.

>>>>

We now evaluate our \esp{} methods on representative real-world datasets of
varying density and label regularity.
After presenting the data and our evaluation criterion, we proceed in two steps. First, we simulate
our generative model on real networks to give them signs, then we study to which extent we can recover
the parameters $p$ and $q$ of each node, and how predictions based on these estimation compare with
the Bayes optimal. Second, we select random training sets from the actual signs.
This shows that our methods compete well against
existing approaches in terms of both predictive and computational performance. We are especially
interested in small training set regimes, and have restricted our comparison to the batch learning
scenario since all competing methods we are aware of have been developed in that setting only.

@@

We consider six real-world classification datasets. The first four are \dssn{} widely
used as benchmarks for this
task~(\eg{}\autocites{Leskovec2010}{shahriari2014ranking}{wu2016troll}{OnlineCompletion17}):
\begin{newcontent}
In \adv{}, a trust-based social network for open source developers, a user $u$ can certify another
user $v$ with different degrees of trust: \enquote{Observer}, \enquote{Apprentice} (both of which we
consider negative), \enquote{Journeyer} and \enquote{Master} (both of which we consider
positive).\footnote{We download the 7\thup{} of July, 2014 version from
\url{http://www.trustlet.org/datasets/advogato/}.} A full description of this trust metric, and its
resistance to attacks, is available in the PhD thesis of the website's creator~\autocite[Section
4]{AdvogatoTrustThesis02}).
\end{newcontent}
In \wik{}, there is an edge from user $u$ to user $v$ if $v$ applies for an admin position and $u$
votes for or against that promotion. In \sla{}, a news sharing and commenting website, member $u$
can tag other members $v$ as friends or foes. Finally, in \epi{}, an online shopping website, user
$v$ reviews products and, based on these reviews, another user $u$ can display whether he considers
$v$ to be reliable or not. In addition to these four datasets, we considered two other \ssn{} where
the signs are inferred automatically, rather than given explicitly by the users.  In
\kiw{}~\cite{wikiedits11}, an edge from Wikipedia user $u$ to user $v$ indicates whether they edited
the same article in a constructive manner or not.\footnote{This is the
\href{http://konect.uni-koblenz.de/networks/wikisigned-k2}{KONECT version of the
\enquote{Wikisigned} dataset}, from which we removed self-loops.} Finally, in the
\aut{}~\cite{kumar2016structure} network, an author $u$ cites another author $v$ by either endorsing
or criticizing $v$'s work. The edge sign is derived by classifying the citation sentiment with a
technique using a list of positive and negative words; see
\cite{kumar2016structure} for more details.\footnote{We again removed self-loops and merged
multi-edges which are all of the same sign.}

\begin{table}[bt]
  \centering
  \small
  \caption{Dataset properties. The 5\thup{} column gives the fraction of positive labels. The next two
  columns provide two different measures of label regularity, while the last two columns give the
  proportion of reciprocal edges, and among them the fraction with different signs.\label{tab:dataset}}
  \begin{tabular}{lrrrrrrrr}
    \toprule
    Dataset & \thead{$|V|$}       & \thead{$|E|$}       & \thead{$\frac{|E|}{|V|}$} &
    \thead{$\frac{|E^+|}{|E|}$} & \thead{$\frac{\Psi^2_{G''}(Y)}{|E|}$} &
    \thead{$\frac{\Psi_G(Y)}{|E|}$} & \thead{reciprocal\\ edges} &
    \thead{reciprocal\\ disagreement} \\
    \midrule                                                                                                   
    \aut{}  & \np{4831}   & \np{39452}  & 8.2               & 72.3\%              & .076                          & .191                    & 5.1\%      & 27.1\%            \\
    \adv{}  & \np{5417}   & \np{51312}  & 9.5               & 75.1\%              & .061                          & .132                    & 33.6\%     & 28.6\%            \\
    \wik{}  & \np{7114}   & \np{103108} & 14.5              & 78.8\%              & .063                          & .142                    & 5.6\%      & 10.0\%            \\
    \sla{}  & \np{82140}  & \np{549202} & 6.7               & 77.4\%              & .059                          & .143                    & 17.7\%     & 4.0\%             \\
    \epi{}  & \np{131580} & \np{840799} & 6.4               & 85.3\%              & .031                          & .074                    & 30.8\%     & 2.1\%             \\
    \kiw{}  & \np{138587} & \np{740106} & 5.3               & 87.9\%              & .034                          & .086                    & 6.5\%      & 14.6\%            \\
    \bottomrule
  \end{tabular}
\end{table}

\autoref{tab:dataset} summarizes statistics for these datasets. We note that most edge labels are
positive. Hence, test set accuracy is not an appropriate measure of prediction performance. We
instead evaluated our performance using the so-called Matthews Correlation Coefficient
(MCC)~\autocite{MCC00}, defined as
\[
	\mathrm{MCC} = \frac{tp\times tn-fp\times fn}%
        {\sqrt{ (tp + fp) ( tp + fn ) ( tn + fp ) ( tn + fn ) } }% = \pm \sqrt{\frac{\chi^2}{n}}
\]
MCC combines all the four quantities found in a binary confusion matrix ($t$rue $p$ositive, $t$rue
$n$egative, $f$alse $p$ositive and $f$alse $n$egative) into a single metric which ranges from $-1$
(when all predictions are incorrect) to $+1$ (when all predictions are correct) through $0$ (when
predictions are made uniformly at random).

Although the semantics of the edge signs is not the same across these networks, we can see from
\autoref{tab:dataset} that our generative model essentially fits all of them. Specifically,
two columns of the table report the rate of label (ir)regularity, as measured by
$\Psi^2_{G''}(Y)/|E|$ (6\thup{} column) and $\Psi_{G}(Y)/|E|$ (7\thup{} column), where 
\[
\Psi^2_{G''}(Y) = \min_{(\bp,\bq)} \left(f_{E_0}(\bp,\bq) + f_{E\setminus E_0}(\bp,\bq)\right)\,,
\]
$f_{E_0}$ and $f_{E\setminus E_0}$ being the quadratic criterions of \autoref{ss:passive}, viewed as
functions of both $(\bp,\bq)$, and \yuv{} when all labels are known, and $\Psi_{G}(Y)$ is the
label regularity measure adopted in the online setting, as defined in \autoref{s:prel}. It is
reasonable to expect that higher label irregularity corresponds to lower prediction performance.
This trend is in fact confirmed by our experimental findings: whereas \epi{} tends to be easy,
\aut{} tends to be hard, and this holds for all algorithms we tested, even if they do not explicitly
comply with our inductive bias principles. Moreover, $\Psi^2_{G''}(Y)/|E|$ tends to be proportional
to $\Psi_{G}(Y)/|E|$ across datasets, hence confirming the anticipated connection between the two
regularity measures.

\begin{newcontent}
Finally, there is a low fraction of reciprocal edges (\ie{} both $\euv \in E$ and $\evu
\in E$), which is a common mechanism of link formation in directed
networks~\autocites{DirectedReciprocity04}{Reciprocity13}). Moreover, in most cases, such reciprocal edges do not disagree,
\ie{} they have the same sign. In practice, we can use this fact to improve our accuracy at no
additional computational cost: when predicting $\euv \in E_{\mathrm{test}}$, if the reciprocal edge
\evu{} is part of the training set, we set $\yhat_{u,v} = y_{v,u}$. For clarity, when comparing
our methods with existing approaches, we not use that trick. But afterwards, we show in
\autoref{tab:twin} when it can be beneficial.
% more references in  http://www.nature.com/articles/srep02729
\end{newcontent}

@@

Recall that according to our generative model of \autoref{s:gen}, each node $u$ has a parameter $p_u$
governing its sending behavior and each node $v$ has a parameter $q_v$ governing its receiving
behavior, such that the sign of the edge \euv{} is positive with probability $\frac{p_u+q_v}{2} =
\etauv$. Given a topology $G=(V,E)$, we start by assigning a $p$ and $q$ values to each node. If we
want \etauv{} to be uniform over $E$, we have to take into account the out-degree of nodes with at
least one outgoing edge, and likewise the in-degree of nodes with at least one incoming edge. For
that, in the case of $p$, we partition the interval $[0,1]$ into $|V|$ segments of size proportional
to $\dout(u_1),
\dout(u_2), \ldots, \dout(u_{|V|})$. We shuffle these segments and draw a number \uar{} from each of
them, which we set as the $p$ value of the corresponding node. Then, because we want to model real
sign distribution and have more
positive edges, we apply an exponential transformation \[ p \rightarrow \frac{1}{1-e^{-\lambda}}
\left( 1-e^{-\lambda p} \right) \,,\] where we choose $\lambda=3$. We do the same for $q$, then for
every edge \euv{}, we set it positive with probability \etauv{}. An example of the distributions we
obtain for the \wik{} graph is showed on \autoref{fig:troll_synth_distrib}, giving $71.9\%$ of
positive edges.

\vspace{\baselineskip}
\noindent\begin{minipage}{187mm} 
  \begin{minipage}{46mm}
    \centering
    \includegraphics[width=.95\linewidth]{raw/synth_wik_pdense.pdf}
    \captionof{subfigure}{Density of $p$.}
    % \label{fig:figure1}
  \end{minipage}%
  \begin{minipage}{46mm}
    \centering 
    \includegraphics[width=.95\linewidth]{raw/synth_wik_qdense.pdf} 
    \captionof{subfigure}{Density of $q$.} 
    % \label{fig:figure2} 
  \end{minipage} 
  \begin{minipage}{46mm}
    \centering 
    \includegraphics[width=.95\linewidth]{raw/synth_wik_etadense.pdf} 
    \captionof{subfigure}{Density of $\eta$.} 
    % \label{fig:figure2} 
  \end{minipage} 
  \begin{minipage}{46mm}
    \centering 
    \includegraphics[width=.95\linewidth]{raw/synth_wik_signdense.pdf}
    \captionof{subfigure}{Density of $\eta$ for both signs.} 
    % \label{fig:figure2} 
  \end{minipage} 
  \captionof{figure}{Synthetic distributions on \wik{}.} 
  \label{fig:troll_synth_distrib} 
\end{minipage}
\vspace{\baselineskip}

Once the signs are generated, we select a training set \trainset{} \uar{}, and predict
the signs of the remaining edges in the testing set $E\setminus \trainset{}$. The Bayes optimal
predictor classifies an edge \euv{} as positive if $\etauv \geq \shalf$ and negative otherwise. 
We compare it with the algorithm analyzed in \autoref{ss:bayes_approx}, which we call \usrule{}
(Bayes Learning Classifier based on \emph{tr}ollness and \emph{un}trustworthiness). In practice,
\usrule{} proceeds as follow: after computing $\htr(u)$ and $\hun(u)$ on training set $E_0$ for all
$u \in V$ (or setting those values to \shalf{} in case there is no outgoing or incoming edges for
some node), we use the equation \eqref{eq:predictor}\footnote{We reproduce this equation here for
convenience: $\sgn\Big(\big(1-\htr(u)\big) + \big(1-\hun(v)\big) -\tau - \tfrac{1}{2}\Big)$.},
having estimated $\tau$ on $E_0$.

The result of this comparison on the six datasets is showed in \autoref{tab:troll_synth_global}.
For each network, we generated the signs once. Then, for different training size (20\%, 40\% or 80\%
of $E$), we sampled a training set and predict using the knowledge of the true $\eta$ (Bayes
predictor) or its estimated value (\usrule{}).
The MCC of the Bayes predictor is the same on all datasets (around $30$\footnote{Here and in the
following, we multiply all MCC value by 100 to improve readability.}), and so is its accuracy
(around $75\%$). The corresponding values for \usrule{} are close on all datasets (although some are
easier) and the gap naturally decreases as the training size increases. Another interesting
quantity is how close can \usrule{} estimates $\eta$ using $\wh{\eta} = \left(1-\htr(u)\right) +
\left(1-\hun(v)\right) -\tauhat$ as defined in \autoref{ss:bayes_approx}. We compute the mean
absolute error (MAE) between $\eta$ and $\wh{\eta}$ for all testing edges, and then specifically for
the testing edges whose both endpoints have been sampled above a certain threshold in the training
set. As expected, the estimation is more accurate with increasing training set size, and with
increasing number of samples for a given edge.

@@

Finally, we can also look in more details at the role of $p$ and $q$ by building a 2D histogram of
the testing edges \euv{} based on their coordinate $(p_u, q_v)$. For instance on
\autoref{fig:tr_synth_mae}, we see that the MAE between $\eta$ and $\wh{\eta}$ is not uniform over
the $p,q$-space. It seems to decrease along the diagonal from $(0,0)$ to $(1,1)$. Note though that
\autoref{fig:tr_synth_sizemae} suggests this is likely due to the higher number of edges in the top
right corner (reflecting the imbalance toward positive edges), which allows better estimation. 
In this case, we cannot use the MCC because off-diagonal, the Bayes predictor classifies edges
either all positive or all negative, resulting in a division by zero in the definition. We thus
fall back on accuracy and display in \autoref{fig:tr_synth_acc} the difference between the accuracy
of the Bayes predictor and the accuracy of \usrule{}. This time, the anti diagonal where
$\frac{p+q}{2}=\lhalf$ seems to play a special role. Indeed, as showed more clearly in
\autoref{fig:tr_synth_cmpacc}, the gap in accuracy between the Bayes predictor and \usrule{}
increases symmetrically as $\frac{p+q}{2}$ moves away from \shalf{}. The gap also tends to disappear
when $\frac{p+q}{2}=\lhalf$, for in that region, both predictors can only rely on random predictions.

While we show those patterns on \wik{} with a 40\% training size, they are consistent across
datasets and training size.
\end{newcontent}

@@

We perform two other sets of experiments. The first one evaluates the effect of predicting reciprocal
edges by their value if available. As expected, the results in \autoref{tab:twin} demonstrate that
it improves MCC when the network has enough reciprocal edges with low disagreements, like in \sla{} and
\epi{}. It has no effect when there are few reciprocal edges, like in \aut{}, \wik{} and \kiw{}. And it is
detrimental when there are many reciprocal edges that do not agree enough, like in \adv{}, in which
case it is better to rely solely on the learned model.
